### 2021-09-15 plos_latex_template.tex
Rocío: minor comments in the rules

Depending on the field, scientific or professional communities can have very different compositions. Thinking about which subgroups have been historically excluded from your community, and thinking about which subgroups hold most privileges can help you guide the organization of a more inclusive and diverse event. Doing this exercise while you think about the reasons why you want to create a more diverse conference can prevent falling into the mistake of treating diversity as a checklist, or using diversity and inclusion work for image purposes.
%andrea: this is here because we want to avoid readers to read as recipes and treat this less seriously than how it should (and then blaming the recipe). i am not sure of the phrasing or position in the text."genuine collaborator" is in the spirit of co-liberation/co-conspirator
%Yanina: They are going to blame what they want to blame. You can't control what people are going to do and say and we shouldn't write defensively because we can't take up space in our message with that. People who do not want to change will not change because of our paper. I imagine this as a guide and help for those people who do want to do something and want to do their best and do not know what or how to do it. Or they have already done something and want to continue deepening those changes. The rest is not worth spending time and energy.  But perhaps the goals of the paper change in some other discussion and I lost that.
Actively be an ally: if you understand your own privileges (unearned advantages given by society to some people but not all) \cite{}, you can help to work on end oppression in your conference. Privilege is largely invisible to those who have it. Once we understand our own privileges, we can begin addressing them on an individual and institutional basis.
%Natalia: I'll look for a reference for privilege definition.
%marcela: I modified the paragraph a bit, but it needs more work to connect Natalia's and Andrea's ideas. It's just a suggestion, feel free to re-write it again.
%Batool: I tried making some paraphrasing - It's just a simple suggestion, not sure if it deliver the message you mean it.  I also think that we may need to define "Privilege" at the 1st mention in the 1st paragraph in this rule. Something like "Privilege is a set of unearned benefits enjoyed by a group, beyond what’s available to others."

@@

While it is essential to improve representation towards some of the most visible dimensions of human diversity, such as race, gender, and country of origin, building a truly inclusive environment means taking care of all the other aspects of diversity as well. Having consideration of religious practices, setting specific accommodations for breastfeeding and child care, having LGBTQIA+-friendly spaces, creating community-only spaces, enforcing the use of pronouns, and treating gender as a non-binary variable are just some examples of decisions that can make inclusion real. Intersectionality also needs to be taken into account, to address multiple inequalities and barriers to unprivileged people that intersect and interact with each other \cite{crenshaw1989}. Importantly, you can take active steps in creating a more welcoming environment without requiring anyone to disclose personal information.

@@

Nevertheless, having a diverse offer of networking opportunities that can appeal to people with different backgrounds, accessibility needs, and preferences can be challenging, especially if you have the mindset of organizing each activity for the complete pool of attendants. If you think of smaller activities that can reach specific groups, preferably co-lead by community leaders, these sessions can be more productive and successful, and even inclusive, than trying to organize one single activity that can please the whole community. Some examples are the newbies sessions for first-timers, mixers lead by specific subgroups or communities, or leisure activities that reunite subgroups with the same interests: arts, exercise, sports, movies, etc.

Adopting a code of conduct and creating a team to enforce it are key aspects in creating a safe environment during a conference \cite{favaroYourScienceConference2016}. The code of conduct is a document meant to keep the community safe and should state clearly: the unacceptable behaviors, the spaces of the conference in which it applies, the consequences for engaging in unacceptable behavior, and the way to report violations \cite{aurora2019}. 
The code of conduct should be displayed prominently in several spaces of the conference to deter people from unacceptable behavior, which includes microaggression. An efficient code of conduct acts as a protection for the community because the people who are the target of unacceptable behavior tend to be the ones with less power or privilege.

@@

Creating and maintaining such a team and space may seem more challenging than working in homogeneous teams, but the positive outcomes far outweigh any inconveniences. 
Having diverse people in decision-making positions will affect positively all the other aspects of your conference because all the processes will benefit from their input, expertise, and distinct perspectives \cite{hongGroupsDiverseProblem2004}. On the other hand, a diverse team plays an important role at creating a welcoming space because seeing people with similar life experiences occupy public spaces, positions of power, and breaking negative stereotypes (i.e., representation) matters (see Rule \ref{rule_inclusion}). This is specially important for networking spaces, where receiving ideas, direct feedback, and spaces to co-lead from a wide variety of community members can be the key to organize meaningful sessions for all the conferences attendees. This gave a sense of variety and an appeal of “at least one networking session for everyone”.

@@

Alternatively, a conference could have a hybrid format with an in-person and online component. This dual format could allow a group of people to interact face-to-face while providing many others the opportunity to participate remotely. The challenge and requirement for this kind of setting would be to make the online component as relevant as the in-person component and not just a consolation prize to the less privileged in the community \cite{ninerBetterWhomLeveling2021}. %Adithi: After the conference is over, the conference materials and videos can be uploaded on free platforms, so that people who were left out of the main event due to any reason, still feel they were cared for and can go to the resources. Also when the event is online, a more timezone friendly approach could be used, all the events if online can be global friendly, people with dependents need hours to take care and the night to rest, if the conference is overnight, they lose sleep and suffer, instead if the conference could be friendly for all regions on certain days just like useR! 2021.

Take advantage of the online setting for new networking activities that can be inclusive for people with different socialization styles and accessibility needs. Social events in in-person conferences are often the most exclusive, and they are a very common place where you have code of conduct violations. Having them directly translated into their virtual versions does not address the need for creating opportunities for people to meet and bond. An online event is a great opportunity to create spaces for people who do not enjoy in person networking. Some ideas in this line are: offering the option of written chat only, instead of voice or video conversations, opening events with team work like trivia, offer some events that can be passively enjoyed like movies, yoga sessions, or art displays, where attendants can choose to just sit and enjoy without talking, or have a chat channel to comment on their experiences during the session. Respect people’s limits, preferences, and remember that “the usual” does not necessarily work with your community, and that one single networking activity will never serve the whole community. It is worth trying new and varied activities that might work with subgroups of attendees.

@@

Conferences are among the least accessible spaces that people with disabilities may encounter \cite{priceAccessImaginedConstruction2009}.
Ironically, accessibility practices are inclusive not only for people with disabilities but can be beneficial to a broad spectrum of people. For instance, having captions is helpful to deaf and hard-of-hearing people, non-native speakers, and everyone in general.

@@

Most importantly, accessibility practices are not afterthoughts that can be dealt with at the last minute. 
They require time and early decision-making \cite{irishIncreasingParticipationUsing2020}. 
Conversely, inaccessible decisions are hard to course-correct, e.g., when finding out too late that a venue is inaccessible for wheelchairs in an in-person conference. 
On the other hand, it is likely that financial constraints may not allow you to make your conference completely accessible and there will be choices to make (e.g. choosing between captions and interpreters). 
Including people with disabilities, or rather, that are rendered disabled by standard practices in conferences in the organizing team would allow them to take part in the decisions from the beginning, spot right away the inaccessible practices that need to improve (see \textbf{Rule \ref{rule_organizing_team}}), and help make the right choices. As Sasha Costanza-Chock and the Design Justice League say, "center the voices of those who are directly impacted by the outcomes of the design process".

@@

>>>>

Depending on the field, scientific or professional communities can have very different compositions. Thinking about which subgroups have been historically excluded from your community, and thinking about which subgroups hold most privileges can help you guide the organization of a more inclusive and diverse event. Doing this exercise while you think about the reasons why you want to create a more diverse conference can prevent falling into the mistake of treating diversity as a checklist, or using diversity and inclusion work for image purposes.
%andrea: this is here because we want to avoid readers to read as recipes and treat this less seriously than how it should (and then blaming the recipe). i am not sure of the phrasing or position in the text."genuine collaborator" is in the spirit of co-liberation/co-conspirator
%Yanina: They are going to blame what they want to blame. You can't control what people are going to do and say and we shouldn't write defensively because we can't take up space in our message with that. People who do not want to change will not change because of our paper. I imagine this as a guide and help for those people who do want to do something and want to do their best and do not know what or how to do it. Or they have already done something and want to continue deepening those changes. The rest is not worth spending time and energy.  But perhaps the goals of the paper change in some other discussion and I lost that.
Actively be an ally: if you understand your own privileges (unearned advantages given by society to some people but not all) \cite{}, you can help to work on end oppression in your conference. Privilege is largely invisible to those who have it. Once we understand our own privileges, we can begin addressing them on an individual and institutional basis.
%Natalia: I'll look for a reference for privilege definition.
%marcela: I modified the paragraph a bit, but it needs more work to connect Natalia's and Andrea's ideas. It's just a suggestion, feel free to re-write it again.
%Batool: I tried making some paraphrasing - It's just a simple suggestion, not sure if it deliver the message you mean it.  I also think that we may need to define "Privilege" at the 1st mention in the 1st paragraph in this rule. Something like "Privilege is a set of unearned benefits enjoyed by a group, beyond what’s available to others."
% Rocío: The paragraph above has quite some redundancy with the paragraph below.

@@

While it is essential to improve representation towards some of the most visible dimensions of human diversity, such as race, gender, and country of origin, building a truly inclusive environment means taking care of all the other aspects of diversity as well. Having consideration of religious practices, setting specific accommodations for breastfeeding and child care, having LGBTQIA+-friendly spaces, creating community-only spaces, enforcing the use of pronouns, and treating gender as a non-binary variable are just some examples of decisions that can make inclusion real.
% Rocío: I'm not sure if "variable" is the right term for non-binary since this is not a statistical analysis but real life.
Intersectionality also needs to be taken into account, to address multiple inequalities and barriers to unprivileged people that intersect and interact with each other \cite{crenshaw1989}. Importantly, you can take active steps in creating a more welcoming environment without requiring anyone to disclose personal information.

@@

% Rocío: there's a link missing between this networking part and the paragraph above. This can go in a networking box.
Nevertheless, having a diverse offer of networking opportunities that can appeal to people with different backgrounds, accessibility needs, and preferences can be challenging, especially if you have the mindset of organizing each activity for the complete pool of attendants. If you think of smaller activities that can reach specific groups, preferably co-lead by community leaders, these sessions can be more productive and successful, and even inclusive, than trying to organize one single activity that can please the whole community. Some examples are the newbies sessions for first-timers, mixers lead by specific subgroups or communities, or leisure activities that reunite subgroups with the same interests: arts, exercise, sports, movies, etc.

Adopting a code of conduct and creating a team to enforce it are key aspects in creating a safe environment during a conference \cite{favaroYourScienceConference2016}. The code of conduct is a document meant to keep the community safe and should state clearly: the unacceptable behaviors, the spaces of the conference in which it applies, the consequences for engaging in unacceptable behavior, and the way to report violations \cite{aurora2019}. 
The code of conduct should be displayed prominently in several spaces of the conference to deter people from unacceptable behavior, which includes microaggression .
%Rocío: Microagression. I know it's not always evident, but that's a detail I would expect in a CoC, not necessarily here unless explicitly justified.
An efficient code of conduct acts as a protection for the community because the people who are the target of unacceptable behavior tend to be the ones with less power or privilege.

@@

Creating and maintaining such a team and space may seem more challenging than working in homogeneous teams, but the positive outcomes far outweigh any inconveniences. 
Having diverse people in decision-making positions will affect positively all the other aspects of your conference because all the processes will benefit from their input, expertise, and distinct perspectives \cite{hongGroupsDiverseProblem2004}. In addition, a diverse team plays an important role at creating a welcoming space because seeing people with similar life experiences occupy public spaces, positions of power, and breaking negative stereotypes (i.e., representation) matters (see Rule \ref{rule_inclusion}). This is specially important for networking spaces, where receiving ideas, direct feedback, and spaces to co-lead from a wide variety of community members can be the key to organize meaningful sessions for all the conferences attendees. This gave a sense of variety and an appeal of “at least one networking session for everyone”. %Rocío: this networking part could go in a box.

@@

Alternatively, a conference could have a hybrid format with an in-person and online component. This dual format could allow a group of people to interact face-to-face while providing many others the opportunity to participate remotely. The challenge and requirement for this kind of setting would be to make the online component as relevant as the in-person component and not just a consolation prize to the less privileged in the community \cite{ninerBetterWhomLeveling2021}. %Adithi: After the conference is over, the conference materials and videos can be uploaded on free platforms, so that people who were left out of the main event due to any reason, still feel they were cared for and can go to the resources. Also when the event is online, a more timezone friendly approach could be used, all the events if online can be global friendly, people with dependents need hours to take care and the night to rest, if the conference is overnight, they lose sleep and suffer, instead if the conference could be friendly for all regions on certain days just like useR! 2021. 
% Rocío: we need to add that just because there's an online component, it doesn't mean that the in-person part of hybrid conferences can be exclusionary.

Take advantage of the online setting for new networking activities that can be inclusive for people with different socialization styles and accessibility needs. Social events in in-person conferences are often the most exclusive, and they are a very common place where you have code of conduct violations.
% Rocío: I think that we should justify why we say that "social events in in-person conferences are often the most exclusive"
Having them directly translated into their virtual versions does not address the need for creating opportunities for people to meet and bond. An online event is a great opportunity to create spaces for people who do not enjoy in person networking. Some ideas in this line are: offering the option of written chat only, instead of voice or video conversations, opening events with team work like trivia, offer some events that can be passively enjoyed like movies, yoga sessions, or art displays, where attendants can choose to just sit and enjoy without talking, or have a chat channel to comment on their experiences during the session. Respect people’s limits, preferences, and remember that “the usual” does not necessarily work with your community, and that one single networking activity will never serve the whole community. It is worth trying new and varied activities that might work with subgroups of attendees.

@@

% Rocío: Could start with an introduction (or put somewhere else) that online can easily make conferences more inclusive for many people, but not necessarily for people with disabilities, and that they can be easily left behind unless we actively work on their inclusion. 
Conferences are among the least accessible spaces that people with disabilities may encounter \cite{priceAccessImaginedConstruction2009}.
Ironically, accessibility practices are inclusive not only for people with disabilities but can be beneficial to a broad spectrum of people. For instance, having captions is helpful to deaf and hard-of-hearing people, non-native speakers, and everyone in general.

@@

Most importantly, accessibility practices are not afterthoughts that can be dealt with at the last minute. 
They require time and early decision-making \cite{irishIncreasingParticipationUsing2020}. 
Conversely, inaccessible decisions are hard to course-correct, e.g., when finding out too late that a venue is inaccessible for wheelchairs in an in-person conference. 
On the other hand, it is likely that financial constraints may not allow you to make your conference completely accessible and there will be choices to make (e.g. choosing between captions and interpreters). 
Including people with disabilities, or rather, that are rendered disabled by standard practices in conferences in the organizing team would allow them to take part in the decisions from the beginning, spot right away the inaccessible practices that need to improve (see \textbf{Rule \ref{rule_organizing_team}}), and help make the right choices. As Sasha Costanza-Chock and the Design Justice League say, "center the voices of those who are directly impacted by the outcomes of the design process".  %Rocío: maybe slightly rephrase the last sentence since it seems to imply that everybody knows Sasha Costanza-Chock.

@@

% Rocío: Rewrite as a design rule.
